---
title: "README"
author: "Fan Li"
date: "Sunday, August 24, 2014"
output: html_document
---

Course: Practical Machine Learning

#### Background

With the increasing popularity of wearable devices for health/exercies monitoring, there is a demand to improve our analytical capability of the data generated by the devices. Particularly it is desirable to know not only how much of a particular activity people do, but also quantify how well they do it. The objective of this project is to develop a statistical model to predict how effective people perform barbell lift exercise based on data from accelerometers on their body.



```{r}
library(knitr)
library(ggplot2)
library(reshape2)
library(dplyr)
library(caret)
library(randomForest)

setwd("C:/Users/Fan/R/Kaggle/Forest Cover Type Prediction")
options(stringsAsFactors=FALSE)

```

You can also embed plots, for example:

```{r, echo=FALSE}
plot(cars)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
